---
title: "Tour de France Route Analysis"
description: |
  Plotting Tour de France stages onto a single map.
author: "Julian During"
date: "`r Sys.Date()`"
output: distill::distill_article
creative_commons: CC BY
repository_url: https://github.com/duju211/tour_de_france_map
base_url: https://www.datannery.com/posts/tour-de-france-route-analysis/
params:
  interactive: TRUE
editor_options: 
  chunk_output_type: console
bibliography: biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = !params$interactive)

source("libraries.R")

df_manifest <- tar_manifest() |>
  mutate(command = str_glue("{name} <- {command}"))
```

```{r, include=FALSE, code=df_manifest$command[df_manifest$name == "tdf_funs_files"]}
```

```{r, include=FALSE, eval=TRUE}
if (params$interactive)
  tar_load(tdf_funs_files)
```


# Idea

Plot the routes of the last Tour de France editions onto a single map.

# Reproducibility

If you want to reproduce this post, you have to perform the following steps:

# Data

In this post the following libraries are used:

```{r, file="libraries.R", eval=FALSE}
```

Website where the data is available from:

```{r base_url, code=df_manifest$command[df_manifest$name == "base_url"]}
```

Define paths to editions:

```{r editions, code=df_manifest$command[df_manifest$name == "editions"]}
```

In the following section, get all links to so called '.gpx' files. These
file represent the geographic data of the routes from the different editions.
To do this perform the following steps:

* Introduce ourselves to the host to identify the rules defined in the
'robots.txt' file (@cit_polite)
* Scrape all html files resulting from the above mentioned editions. By using
functions from the 'polite' package we automatically obey the rules defined
from the host
* Search for all html elements that represent links (@cit_rvest)
* Further filter those links for elements with the text 'GPX' in them
(excluding links that don't represent '.gpx' files) and extract the 'href'
attribute
* Put them into a final data frame and extract the year of the edition from
the url (@cit_tidyverse)

```{r fun_gpx_paths, file=tdf_funs_files[["R/gpx_paths.R"]]}
```

Define the CSS expression to identify relevant links. This expression was 
determined using the developer tools of the browser:

```{r links_css, code=df_manifest$command[df_manifest$name == "links_css"]}
```

Execute the above described function and make sure that every resulting path
is unique (@cit_assertr):

```{r df_gpx_paths, code=df_manifest$command[df_manifest$name == "df_gpx_paths"]}
```

```{r, include=FALSE, eval=TRUE}
if (params$interactive)
  tar_load(df_gpx_paths)
```

```{r, echo=FALSE, eval=TRUE}
df_gpx_paths
```

Now all the links to the relevant files are available.

The next step is to read in all the files and determine relevant elements from
these files. To to this perform the following steps:

* As in the previous function use 'polite' functions to scrape all the files
* Search for elements that represent so called 'trackpoints'
* Extract longitudinal information from these elements and put them in a
data frame

```{r, file=tdf_funs_files[["R/track_points.R"]]}
```

Before we call the function define the CSS that will select the right elements.
Again this was determined using the SelectorGadget tool:

```{r track_point_css, code=df_manifest$command[df_manifest$name == "track_point_css"]}
```

Insert the base url, the links to the gpx files and the mentioned CSS
selector into the function and call it:

```{r df_track_points, code=df_manifest$command[df_manifest$name == "df_track_points"]}
```

```{r, include=FALSE, eval=TRUE}
if (params$interactive)
  tar_load(df_track_points)
```

```{r, echo=FALSE, eval=TRUE}
df_track_points
```

To ease the use of spatial data, turn the data frame into a 'sf'
(@cit_sf) object:

```{r, file=tdf_funs_files[["R/stages_sf.R"]]}
```

If we apply the function, we get an data frame with one row per stage and
a geometry list column, that represents the spatial data of that stage:

```{r sf_tdf_stages, code=df_manifest$command[df_manifest$name == "sf_tdf_stages"]}
```

```{r, eval=TRUE, include=FALSE}
if (params$interactive)
  tar_load(sf_tdf_stages)
```

```{r, eval=TRUE, echo=FALSE}
sf_tdf_stages
```

Before we plot the data download the outlines of countries with the help of
the 'rnaturalearth' (@cit_rnaturalearth) package.

```{r world, code=df_manifest$command[df_manifest$name == "world"]}
```

Combine the stage data with the country outline data and plot the result using
ggplot techniques:

```{r, file="R/vis_tdf_stages.R"}
```

```{r gg_tdf_stages, code=df_manifest$command[df_manifest$name == "gg_tdf_stages"]}
```

```{r, include=FALSE, eval=TRUE}
if (params$interactive)
  tar_load(gg_tdf_stages)
```

```{r plot_tdf_stages, echo=FALSE, eval=TRUE, preview=TRUE}
gg_tdf_stages
```


